---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data
```{r 1.1}
data<-read.csv("activity.csv")
str(data)
```


2. Process/tranform the data into a format suitable for your analysis
```{r 1.2}
data$date<-as.Date(data$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day
```{r 2.1}
date<-unique(data$date,fromLast = TRUE)
tstep<-vector(mode="numeric")
for(i in date){
       tstep<-c(tstep,sum(data[data$date==i,]$steps,na.rm=TRUE))
}
plot(date,tstep,xlab="Date",ylab="Total number of steps",main="The total number of steps taken each day",type="h",lwd=6,col="blue")
```


2. Calculate and report the mean and median total number of steps taken per day

Here's the **mean** total number of steps taken per day
```{r 2.2.1}
meanstep<-vector(mode="numeric")
for(i in date){
       meanstep<-c(meanstep,mean(data[data$date==i,]$steps,na.rm=TRUE))
}
print(data.frame(Date=date,Mean.Step=meanstep))
```

Here's the **median** total number of steps taken per day
```{r 2.2.2}
medstep<-vector(mode="numeric")
for(i in date){
       medstep<-c(medstep,median(sort(data[data$date==i,]$steps)))
}
print(data.frame(Date=date,Median.Step=medstep))
```


## What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r 3.1}
itv<-unique(data$interval,fromLast = TRUE)
avgstep<-vector(mode="numeric")
for(i in itv){
       avgstep<-c(avgstep,mean(data[data$interval==i,]$steps,na.rm=TRUE))
}
meaninterval<-data.frame(itv,avgstep)
plot(itv,avgstep,xlab="5-minute interval",ylab="Average number of steps",main="Time series plot of 5-minute interval",type="l",lwd=1,col="red")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r 3.2}
itv[which.max(avgstep)]
```


## Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r 4.1}
sum(is.na(data$steps))
sum(is.na(data$date))
sum(is.na(data$interval))
sum(is.na(data))
```

2. Filling in all of the missing values in the dataset by mean for that 5-minute interval
```{r 4.2}
newdata<-data
a<-which(is.na(newdata$steps))
for(i in a){
       newdata$steps[i]<-meaninterval[meaninterval$itv==newdata$interval[i],]$avgstep
}
```

3. Creat a new data
```{r 4.3}
str(newdata)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

Here's a histogram of total number of steps taken each day in new data
```{r 4.4.1}
ntstep<-vector(mode="numeric")
for(i in date){
       ntstep<-c(ntstep,sum(newdata[newdata$date==i,]$steps,na.rm=TRUE))
}
plot(date,ntstep,xlab="Date",ylab="Total number of steps",main="The total number of steps taken each day",type="h",lwd=6,col="blue")
```

Here's the **mean** total number of steps taken each day
```{r 4.4.2}
nmeanstep<-vector(mode="numeric")
for(i in date){
       nmeanstep<-c(nmeanstep,mean(newdata[newdata$date==i,]$steps,na.rm=TRUE))
}
print(data.frame(Date=date,Mean.Step=nmeanstep))
```

Here's the **median** total number of steps taken each day
```{r 4.4.3}
nmedstep<-vector(mode="numeric")
for(i in date){
       nmedstep<-c(nmedstep,median(sort(newdata[newdata$date==i,]$steps)))
}
print(data.frame(Date=date,Median.Step=nmedstep))
```

First, we find the day had NAs filled
```{r 4.4.4}
unique(data$date[which(is.na(data$steps))])
```
So, the impact of imputing missing data on the estimates of the total daily number of steps is on that day.


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r 5.1}
wd<-vector(mode="character")
nr<-nrow(newdata)
for (i in 1:nr){
       ifelse(weekdays(newdata$date[i])!="Saturday"&weekdays(newdata$date[i])!="Sunday",wd<-c(wd,"weekday"),wd<-c(wd,"weekend"))
}
newdata<-data.frame(newdata,wd)
str(newdata)
```

2. Make a panel plot containing a time series plot of the 5-minute interval
```{r 5.2}
library(lattice)
numstep<-vector(mode="numeric")
for(i in itv){
       for (j in c("weekday","weekend")){
              numstep<-c(numstep,mean(newdata[newdata$interval==i&newdata$wd==j,]$steps))
       }
}
d<-data.frame(interval=rep(itv,each=2),wd=rep(c("weekday","weekend"),times=length(itv)),numstep)
xyplot(numstep~itv|wd,data=d,layout=c(1,2),xlab="Interval",ylab="Number of steps",type="l",lwd=1,col="blue")
```